This is the OSF repo: https://osf.io/fes78/overview

```{r include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(osfr)
library(jsonlite)
library(purrr)
```

```{r include=FALSE}
osf_retrieve_node("fes78") |>
  osf_ls_files(pattern=".json", n_max=Inf) |>
  osf_download()
```

```{r include=FALSE}
data <- list.files(pattern=".json") |>
  lapply(fromJSON) |>
  bind_rows()
```

```{r include=FALSE}
visual_search_data <- data |>
  filter(task=="visual_search", is_practice == FALSE) |>
  select(subject, trial_index, set_size, search_type, target_present, rt, correct, tobii_data, item_positions)
```

```{r include=FALSE}
visual_search_summary_data <- visual_search_data |>
  group_by(subject, set_size, search_type, target_present) |>
  summarize(M=mean(rt))
```

```{r echo=FALSE}
ggplot(visual_search_summary_data, aes(color=target_present, x=set_size, y=M)) +
  facet_wrap("search_type") +
  geom_point() +
  geom_smooth(method="lm")
```


```{r include=FALSE}
eye_tracking_data <- visual_search_data |>
  #unnest_wider(item_positions, names_sep="_") |>
  unnest(tobii_data) |>
  select(-leftPupilDiameter, -rightPupilDiameter) |>
  mutate(x=1920*x, y=1080*y)
```



Saccade detection

```{r include=FALSE}
compute_fixations <- function(data) {

  sampling_interval <- median(diff(data$timestamp))

  d <- data |>
    mutate(vx = (lead(x, 2) + lead(x, 1) - lag(x, 1) - lag(x, 2))/(6*sampling_interval)) |>
    mutate(vy = (lead(y, 2) + lead(y, 1) - lag(y, 1) - lag(y, 2))/(6*sampling_interval)) |>
    filter(!is.na(vx), !is.na(vy))

  sigma_x <- sqrt(median(d$vx^2) - median(d$vx)^2)
  sigma_y <- sqrt(median(d$vy^2) - median(d$vy)^2)

  lambda <- 6

  thresh_x <- lambda * sigma_x
  thresh_y <- lambda * sigma_y

  d <- d |>
    mutate(is_saccade = (vx / thresh_x)^2 + (vy / thresh_y)^2 > 1) |>
    mutate(event = cumsum(is_saccade != lag(is_saccade, default=first(is_saccade))))

  fixations <- d |>
    group_by(event) |>
    summarize(
      is_saccade = first(is_saccade),
      x = median(x),
      y = median(y),
      start = first(timestamp),
      end = last(timestamp),
      duration = last(timestamp) - first(timestamp),
      n_samples = n()
    ) %>%
    filter(!is_saccade)

  return(fixations)
}
```

# Extract data frame of item positions
```{r}
item_locations <- eye_tracking_data|>
  select(subject, trial_index, search_type, set_size, target_present, item_positions) |>
  group_by(subject, trial_index, search_type, set_size, target_present) |>
  slice_head(n=1) |>
  unnest(item_positions)
```

```{r}

one_trial <- eye_tracking_data |>
  filter(subject=="3t1hlagk9xf9", trial_index==86)

stimuli <- (one_trial |> slice_head(n=1) |> select(item_positions))[[1]][[1]]

fixations <- compute_fixations(one_trial)

ggplot(stimuli, aes(x=x, y=y)) +
  geom_point(size=5, mapping=aes(color=color, shape=shape)) +
  geom_point(data=one_trial, size=1, shape="circle", color="grey50") +
  geom_point(data=fixations, size=10, shape="circle open", color="black") +
  geom_path(data=fixations, color="black", arrow=arrow())+
  coord_cartesian(xlim=c(0,1920), ylim=c(0, 1080)) +
  scale_y_reverse() + 
  scale_color_manual(values=c("#0D47A1", "#E53935"), guide=FALSE)+
  scale_shape_manual(values=c("circle open", "cross"), guide=FALSE)+
  theme_void()+
  theme(panel.border = element_rect(color="black", linewidth=1))

```

# Extract fixations per trial

```{r include=FALSE}
fixations <- eye_tracking_data |>
  group_by(subject, trial_index, search_type, set_size, target_present) |>
  reframe(compute_fixations(pick(everything())))
```


# Count fixations for each trial

```{r}
fixation_counts <- fixations |>
  group_by(subject, trial_index, search_type, set_size, target_present) |>
  summarize(n_fixations = n())
```

```{r echo=FALSE}
ggplot(fixation_counts, aes(x=set_size, y=n_fixations, color=search_type)) +
  geom_point() +
  facet_wrap("target_present") +
  geom_smooth(method="lm")
```

# FVF estimates

Algorithm is:
1. create radius around each fixation point in target present trials.
2. expand radius until it contains 50% of all items on the screen.
3. that is FVF for that trial


```{r include=FALSE}
calculate_coverage <- function(fixations_trial, items_trial, radius) {

  # For each distractor, check if ANY fixation is within radius
  # Cross join: every distractor × every fixation
  visited <- crossing(
    items_trial %>% select(item_x = x, item_y = y),
    fixations_trial %>% select(fix_x = x, fix_y = y)
  ) %>%
    mutate(distance = sqrt((item_x - fix_x)^2 + (item_y - fix_y)^2)) %>%
    group_by(item_x, item_y) %>%
    summarise(visited = any(distance <= radius), .groups = "drop")
  
  mean(visited$visited)
}

find_fvf <- function(fixations_trial, items_trial, 
                     radii = seq(10, 400, by = 5)) {
  
  coverages <- sapply(radii, function(r) {
    calculate_coverage(fixations_trial, items_trial, r)
  })
  
  # Find first radius where coverage >= 0.5
  threshold_idx <- which(coverages >= 0.5)[1]
  
  if (is.na(threshold_idx)) {
    return(list(fvf = NA, radii = radii, coverages = coverages))
  }
  
  list(
    fvf = radii[threshold_idx],
    radii = radii,
    coverages = coverages
  )
}

```


```{r include=FALSE}
nested_items <- item_locations |>
  group_by(subject, trial_index, search_type, set_size, target_present) |>
  nest() |>
  rename(item_locations = data)

nested_fixations <- fixations |>
  group_by(subject, trial_index, search_type, set_size, target_present) |>
  nest() |>
  rename(fixations = data)

fvf_data <- nested_items |>
  left_join(nested_fixations, by=c("subject", "trial_index", "search_type", "set_size", "target_present"))
```

```{r include=FALSE}
fvf_result <- fvf_data |>
  filter(target_present == FALSE) |>
  mutate(fvf = map2(fixations, item_locations, find_fvf)) %>%
  unnest_wider(fvf) |>
  select(-radii, -coverages, -item_locations, -fixations)
```

```{r echo=FALSE}
ggplot(fvf_result, aes(x=set_size, y=fvf, color=search_type))+
  facet_wrap("search_type")+
  geom_point()+
  geom_smooth(method="loess")
```





# Notes on things to fix in the data

1. there's no pixel-level eye tracking. it is in normalized coords.
2. three timestamps feels excessive. figure out which one should be included.